---
title: "Distributions"
author: "Brian"
date: "15 August 2015"
output: html_document
---

<!--
echo = show the code
results = 'asis' or 'hide' or 'hold' - hold till the end
messages: option message (FALSE hides messages in the output)
warnings: option warning (FALSE hides warnings in the output)
errors: option error (FALSE will make R stop if an error occurs; TRUE will show the error messages in the output)
include = show the code-->

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=3, fig.align ='center',
                      fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE,
                      results='hold',
                      include=TRUE,
                      cache=TRUE)

# If you run into problems with cached output you can always clear the knitr cache by removing the folder named with a _cache suffix within your document’s directory.

```

# r,d,q,p in R

R comes with a rich set of probability distributions, and four consistent ways of accessing them: r, d, q, and p. If you combine these with the name of the distribution - rnorm(), dnorm(), qnorm(), and pnorm(), for example - you can generate random numbers, find the value of the function at a specific point, produce a critical value, and calculate a p-value. 

### r
When you want to generate random numbers, use the r prefix, as in rf() for the F distribution. Most distributions require one or more parameters to define the shape of the distribution. For the F distribution, you need the degrees of freedom for the numerator (df1) and the denominator (df2). To generate n=50 random numbers;

```{r,echo=TRUE}
rf(n=50, df1=15, df2=20)
hist(rf(100000, df1=15, df2=20), breaks=100, col="gray", xlab="F", main="")
```

###d

The d prefix, such as df(), is useful for finding the value of the F function, called its density, at any value of the statistic. Useful to look at the shape o fa curve also from a sequence.

```{r,echo=TRUE}
df(x=2.3, df1=15, df2=20)
myF <- seq(from=0, to=6, by=0.01)
dF <- df(x=myF, df1=15, df2=20)
plot(myF, dF, type="l", lwd=3, col="blue")
```

### q

The q prefix, as in qf(), is how you obtain critical values from a probability distribution. By default, R will integrate the probability starting from the left tail. If you want to get the probability starting from the right tail, be sure to set lower.tail=FALSE. 
```{r,echo=TRUE}
alpha = 0.05
qf(p=alpha, df1=15, df2=20, lower.tail=FALSE)
```

p
The p prefix, as in pf(), is how you calculate a p-value from a probability distribution. As for the q prefix, R will always integrate starting with the lower tail. If you want the right tail, specify lower.tail=FALSE. Specify the value of F for which you want a p-value, the degrees of freedom, and whether you want the left tail or the right tail.

```{r}
observedF = 3.7
pf(q=observedF, , df1=15, df2=20, lower.tail=FALSE)
```

# Normal Distribution

The normal distribution appears in many biological and other process. It is symetrical and is formed by the *central limit theorem* Many random effects will be approximately normally distributed. So peoples height, weight etc. 

```{r norm1, fig.show='hold',,fig.width=10,echo=FALSE}
library(UsingR)
library(ggplot2)
data<-babies

#Create two plots one of the actual data, and one transpose to the Standard Normal N(0,1)
#Using ggplot
norm1a <- ggplot(data, aes(x=wt)) +
  geom_histogram(colour = "darkgreen", fill = "white",aes(y = ..density..)) +
  geom_density(colour="darkgreen")+geom_density(adjust=5,colour="red") +
  xlab('Babies Weight')+ylab('density') 


norm1a <- ggplot(data, aes(x=wt)) +
  geom_histogram(colour = "darkgreen", fill = "white",aes(y = ..density..)) +
  geom_density(colour="darkgreen")+geom_density(adjust=5,colour="red") +
  xlab('Babies Weight')+ylab('density') +
  ggtitle("Histogram of Babies Weights") 



data$wtscale<-scale(data$wt,center=TRUE,scale=TRUE)

#z score density - there probably is some function for this? 
zdenz <- function (x) {
 zd<-(1/sqrt(2*pi))*exp(-1*((x^2)/2))
 return(zd)}

norm1b<-ggplot(data, aes(x=wtscale)) +
  geom_histogram(colour = "darkgreen", fill = "white",aes(y = ..density..)) +
  geom_density(colour="darkgreen")+geom_density(adjust=5,colour="red") +
  xlab('Babies Weight')+ylab('density') +
  ggtitle("Histogram Scaled N(0,1) Babies Weights \n Z density is calculated") +
  geom_segment(aes(x = -1, y = 0, xend = -1, yend = zdenz(-1)),lty=2,colour="blue")+
  geom_segment(aes(x = 1, y = 0, xend = 1, yend = zdenz(1)),lty=2,colour="blue")+
   geom_segment(aes(x = -2, y = 0, xend = -2, yend = zdenz(-2)),lty=2,lwd=2,colour="black")+
  geom_segment(aes(x = 2, y = 0, xend = 2, yend = zdenz(2)),lty=2,lwd=2,colour="black")

library(gridExtra)
grid.arrange(norm1a, norm1b, ncol=2)



```

#Chi Squared Distribution

If you sample from a normal distribution and sqaure the results you will get a $\chi^{2}$ distribution. More formally $X \sim \aleph(0,1)$. Depending on the number of samples that are taken, the degrees of freedom pushes the probablity density distribution to the right until it forms a normal distribution itself.  

The distribution is additive. Therefore $\chi^2_1=X^2_1 \:and\: \chi^2_2=X^2_1+X^2_2$ etc. 


```{r chi1, echo=FALSE,fig.show='hold'}

#lets do in ggplot


x <- rchisq(100, 5)
hist(x,prob=TRUE,xlim=c(0,35),main="Chi Square Distribution")
curve( dchisq(x, df=1), col='green', add=TRUE)
curve( dchisq(x, df=2), col='red', add=TRUE)
curve( dchisq(x, df=3), col='blue', add=TRUE)
curve( dchisq(x, df=4), col='black', add=TRUE)
curve( dchisq(x, df=5), col='darkgreen', add=TRUE)
curve( dchisq(x, df=10), col='red', add=TRUE )


#Function to sample from data Lets sample from
chisamp<-function(degfree){
take <- sample(unique(data$wtscale), degfree)
take<-take^2
take<-sum(take)
return(take)}


#data frame for the results
chisampdf <- data.frame(z= numeric(0),degf= character(0))
for (i in 1:5 ) {
  results <- sapply(1:1000, function(x) { chisamp(i) + 1}) #apply the function chisamp 1000 times 
  tempdf<-data.frame(z=results,degf=paste("df",i))
  chisampdf<-rbind(chisampdf,tempdf)
}  
  
chi1a<-ggplot(chisampdf, aes(z, fill = degf)) + geom_density(alpha = 0.2,adjust=5)
chi1a<-chi1a+ggtitle("Chi Square Distribution of Babies Weight, smoothed")
chi1a

```

#F Distribtuion

If V 1 and V 2 are two independent random variables having the Chi-Squared distribution with m1 and m2 degrees of freedom respectively, then the following quantity follows an F distribution with m1 numerator degrees of freedom and m2 denominator degrees of freedom, i.e., (m1,m2) degrees of freedom.

$$F = \frac{V1_{m1}}{V2_{m2}} \sim F_{(m1,m2)}$$

Now on to the t-distribution and the F-distrution ratio test. 

F and chi-squared statistics are really the same thing in that, after a normalization, chi-squared is the limiting distribution of the F as the denominator degrees of freedom goes to infinity. The normalization is chi-squared = (numerator degrees of freedom) * F. 

```{r}

qf(.95, df1=5, df2=2) 


#F Distribution - right probability
1-pf(2.05,df1=2,df2=71)
1-pf(2.05,df1=2,df2=150)
1-pf(2.05,df1=2,df2=1500)
1-pf(2.05,df1=2,df2=15000)

#Significant value 
qf(0.95,df1=2,df2=71)
qf(0.95,df1=2,df2=15000)


1-pchisq(4.05, df=2) 

qchisq(0.1287,df=2)

1-pchisq(3,2)





```



